---
###title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
  1.  Loading the Data
```{r}
rm(list=ls())
library(data.table)
table=data.table(read.csv("activity.csv"))
```

  2.  Processing the Data
```{r}
complete=table[complete.cases(table)] # table without NA
stepsbyday=complete[,.(total=sum(steps)),     
                    #Agregate 
			              by = .(date)]
stepsbyinterval=complete[,.("Average Daily steps"=mean(steps)), 
                         #Agregate 
                         by = .(interval)]
```

## What is mean total number of steps taken per day?

  1. Calculate the total number of steps taken per day:
```{r}
steps=stepsbyday$total
```

  2. Construct Histogram with calculated data:
```{r} 
hist(steps,col = "blue",main="Histogram of the total number of steps taken per day")
```

  3.  Calculate and report the mean:

The mean total number of steps taken is `r as.integer(mean(steps))` and 
the median is `r median(steps)`


## What is the average daily activity pattern?
  1.   Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
#Plotting the Average Daily steps as a function of 5-min Interval
with(stepsbyinterval,plot(interval,`Average Daily steps`,type="l",xlab="5-min Interval",main="Average # of Daily steps vs 5-min Interval"))
```

  2.   Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
#Get max interval value 
MaxInterval= stepsbyinterval[which.max(`Average Daily steps`)]$interval
```
The interval with maximum average number of steps is `r MaxInterval`.

## Imputing missing values
  1.  Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
NumberOfRowsWithNA=dim(table)[1]-dim(complete)[1]
```
There is  `r NumberOfRowsWithNA` rows with missing values.

  2.  Devise a strategy for filling in all of the missing values in the dataset. 
```{r}
	#Deciding Strategies
	isThereEnoughInformationToFillNAwithDailyMeanSteps=length(intersect(table[is.na(steps)]$date,stepsbyday$date))>0

#is There Enough Information To Fill NA with Daily Mean Steps?
  #Answer == FALSE/NO
```

The chosen strategy is to fill the **NA**(s) with the  mean for that 5-minute interval.

  3.  Create a new dataset that is equal to the original dataset but with the missing data filled in.
  
```{r}
#Lets fill the Na(s) with the 5-min-average for all days:
	fillNa=table[,.(imean=mean(steps,na.rm=TRUE),steps,date),
			by =.(interval)]
	fillNa[is.na(steps)]$steps=as.integer(fillNa[is.na(steps)]$imean)

#Extracting total steps taken with filled table
	fillSteps=fillNa[,.(total=sum(steps)),
			by = .(date)]$total
```

  4.  Make a histogram of the total number of steps taken each day  
  
```{r}
#Ploting
	hist(fillSteps,col="blue",main = "Total number of steps/day (imputed data) vs  5-min Interval ")
```
 	
 	and Calculate and report the mean and median total number of steps taken per day.
 	
The mean total number of steps taken for the imputed data is `r as.integer(mean(fillSteps))` and 
the median is `r median(fillSteps)`
	

## Are there differences in activity patterns between weekdays and weekends?

  1.  Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
#Add Categories
# To add categories I calculated the number of days since the first date of the dataset.
# Knowning it to be Monday it was just a matter to divide by 7 and taking the rest.
# the rest+1 is in the range 1:7, dividing by 6 and rounding down I have 0 associated with
# weekdays and 1 associated with weekends.
# This number is a possible index to the array below which produce the desired category.
#( I chose this method because it is Locale insensitive.)
	WeekCategories=c("weekday","weekend")
	fillNa[,day:=as.factor(
			WeekCategories[
				as.integer(
				(as.numeric(as.Date(date)-as.Date("2012-10-01"))%%7+1)/6)+1])]
```

	2.  Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). :
```{r}
	WeekDay=fillNa[day==WeekCategories[1],
			.("Number of steps"=mean(steps))
                     	,by = .(interval)]
  	WeekDay$day=WeekCategories[1]
  	WeekEnd=fillNa[day==WeekCategories[2],
			.("Number of steps"=mean(steps))
			,by = .(interval)]
	WeekEnd$day=WeekCategories[2]

#Putting it all back together  
	week=merge(WeekDay,WeekEnd,all = TRUE ,by = names(WeekDay))

#We plot
	library(lattice)
	xyplot(`Number of steps` ~  interval | day,data=week,layout= c(1,2),type="l",xlab="Interval")
```

